---
title: "Report assignment"
author: "Isabelle Wethli"
format: 
 html:
  toc: true
  code-fold: show
  code-tools: true 
  df-print: paged 
  embeded-resources: true 
---

## Introduction

The aim is to visualize the effects of smoking on insurance charges.

## Data

```{r}
library(tidyverse)
library(ggplot2)
library(gtsummary)
library(here)

dat <- read_csv(here("data/raw/insurance_with_date.csv"))
str(dat)

dataselect <- dat |>
  select(sex, bmi, smoker, charges, region) |>
  mutate(sex = factor(sex, levels=c("male", "female"), labels=c("m","f")),
         smoker = factor(smoker))
```

## Descriptive Statistics

This helps to visualize the effect of smoking on insurance charges.

```{r}
tabl_summary <- dat |>   
  tbl_summary(
    by = smoker,
    type = all_continuous() ~ "continuous2",
    include =c("charges","bmi","age"),
    statistic = all_continuous() ~ c("{mean} ({sd})", 
                                     "{median} ({p25}, {p75})", 
                                     "{min}, {max}")
  )
tabl_summary
```

## 

Boxplot 1: Smokers vs. Non-Smokers - Who has more charges?

```{r}
ggplot(dataselect, aes(x=smoker, y=charges, fill = smoker)) + 
  geom_boxplot() +
  ggtitle(label = "Smokers vs. Non-Smokers") +
  xlab(label = "") +
  ylab(label = "Charges from insurance") +
  theme_bw() + theme(legend.position="bottom") 
```

Histogram of charges

```{r}
ggplot(dat, aes(x=charges)) + 
  geom_histogram(bins=50)
```

This histogram shows that the data is not normalized. In order to do a T-Test on the data, we have to logarithmise the charges.

## Inferential Statistics

Logarithmise the charges

```{r}
#| output: false 
  dat$chargeslog <- log(dat$charges)
dat$chargeslog
```

Histogram of charges

```{r}
ggplot(dat, aes(x=chargeslog, fill = smoker)) + 
  geom_histogram(bins=50) +
  theme_bw() + theme(legend.position="bottom") 
```

Test normality

```{r}
dat %>%
  ggplot(aes(sample = chargeslog, fill = smoker, colour = smoker)) +
  geom_qq_line(distribution =  stats::qnorm) +
  geom_qq(distribution =  stats::qnorm) +
  xlab("theoretical") +
  ylab("sample") +
  theme_classic() +
  facet_wrap("sex")
```

## T-Test charges smokers vs. non-smokers

H0: Smoking is not associated with the insurance charges 
HA: Smoking is associated with insurance charges

```{r}
t.test(chargeslog ~ smoker, data = dat)
```
